Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-06-19 ms.tex
Typo
Fix marginpar formatting
Added description of Gaussian quadratures breakpoints

Ground-based \ac{GW} interferometers are entering the advanced detector era, positioning themselves to make profound discoveries about the universe. The \ac{aLIGO} and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India. Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$~\cite{rates}.%
%
\marginpar{Check that number.}
%
In addition to being an efficient source of \acp{GW}, the \ac{NS} may be tidally shredded before merging, providing fuel for an electromagnetic counterpart. The short-lived accretion flow may power a collimated relativistic jet, resulting in a short \ac{GRB} and an X-ray afterglow if it is aligned with the line of sight. An optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium. As surrounding neutron-rich ejecta decay radioactively, an optical `kilonova' may be visible after $\sim$1 day \cite{metzger:2010}. Finally, as the ejecta plow into the interstellar medium, a faint radio afterglow could occur on a timescale of weeks to years \cite{Nakar:2011cw}.

@@

We have no explicit prior on the intrinsic paramters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}$. The distance prior is given by a user-selected power of distance, ${D_\mathrm{L}}^m$, where $m=2$ for a prior that is uniform in volume, and $m=-1$ for a scale-invariant prior that is uniform in $\log D_\mathrm{L}$. The priors are bounded as follows:
%
\marginpar{Need to discuss $D_{\mathrm{L},\mathrm{min}}$, $D_{\mathrm{L},\mathrm{max}}$, and how they are selected automatically if the user does not provide them.}
%
\begin{eqnarray*}
	D_\mathrm{L} &\in& [D_{\mathrm{L},\mathrm{min}}, D_{\mathrm{L},\mathrm{max}}] \\
	t_\oplus &\in& [-\infty, \infty] \\
	u &\in& [0, 1] \\
	\psi &\in& [0, \pi/2)
\end{eqnarray*}
%
The prior does not permit negative values of $u$ because only even powers of $u$ appear in \Eref{eq:snr-model}. The prior only permits $\psi$ to run to $pi/2$ because only trigonometric functions of $4\psi$ appear in \Eref{eq:snr-model}.

@@

>>>>

Ground-based \ac{GW} interferometers are entering the advanced detector era, positioning themselves to make profound discoveries about the universe. The \ac{aLIGO} and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India. Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$~\cite{rates}.%
%
\marginpar{Check that number.}%
%
In addition to being an efficient source of \acp{GW}, the \ac{NS} may be tidally shredded before merging, providing fuel for an electromagnetic counterpart. The short-lived accretion flow may power a collimated relativistic jet, resulting in a short \ac{GRB} and an X-ray afterglow if it is aligned with the line of sight. An optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium. As surrounding neutron-rich ejecta decay radioactively, an optical `kilonova' may be visible after $\sim$1 day \cite{metzger:2010}. Finally, as the ejecta plow into the interstellar medium, a faint radio afterglow could occur on a timescale of weeks to years \cite{Nakar:2011cw}.

@@

We have no explicit prior on the intrinsic paramters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}$. The distance prior is given by a user-selected power of distance, ${D_\mathrm{L}}^m$, where $m=2$ for a prior that is uniform in volume, and $m=-1$ for a scale-invariant prior that is uniform in $\log D_\mathrm{L}$. The priors are bounded as follows:
%
\marginpar{Need to discuss $D_{\mathrm{L},\mathrm{min}}$, $D_{\mathrm{L},\mathrm{max}}$, and how they are selected automatically if the user does not provide them.}
%
\begin{eqnarray*}
	D_\mathrm{L} &\in& [D_{\mathrm{L},\mathrm{min}}, D_{\mathrm{L},\mathrm{max}}] \\
	t_\oplus &\in& [-\infty, \infty] \\
	u &\in& [0, 1] \\
	\psi &\in& [0, \pi/2)
\end{eqnarray*}
%
The prior does not permit negative values of $u$ because only even powers of $u$ appear in \Eref{eq:snr-model}. The prior only permits $\psi$ to run to $\pi/2$ because only trigonometric functions of $4\psi$ appear in \Eref{eq:snr-model}.

@@

Before evaluating each inner integral over $D_\mathrm{L}$, the quantities $A=-\frac{1}{2} \sum_i {\rho_i}^2$ and $B=\sum_i \rho_i \hat\rho_i$ are precalculated so that the integrand can be written as
%
\begin{equation*}
	\exp(A {D_\mathrm{L}}^{-2} + B D_\mathrm{L}) {D_\mathrm{L}}^m.
\end{equation*}
%
Then, we see that the likelihood $\exp(A {D_\mathrm{L}}^{-2} + B D_\mathrm{L})$ is maximized when $D_\mathrm{L} = D_{\mathrm{L},0} = -2A/B$. The likelihood takes on a factor $\eta$ of its maximum value when
%
\begin{equation}
	D_\mathrm{L} = D_{\mathrm{L},\pm} = \left(\frac{1}{D_{\mathrm{L},0}} \pm \sqrt{\frac{\log\eta}{A}}\right)^{-1}.
\end{equation}
